---
output: html_document
---
Olive MGK
Nov, 2018
Udacity Data Analyst Nanodegree
========================================================

## Exploring and Summarizing White Wine Data with R

Project Overview:  The dataset contains white variants of the Portuguese "Vinho 
Verde" wine.  There are 4898 white wine instances, 11 + output number of 
Attributes.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

install.packages('RJSONIO', dependencies = TRUE, repos = 
                   "http://cran.us.r-project.org")

install.packages('ggplot2', dependencies = TRUE, repos = 
                   "http://cran.us.r-project.org")
install.packages('RColorBrewer', dependencies = TRUE, repos = 
                   "http://cran.us.r-project.org")
install.packages('gridExtra',repos = "http://cran.us.r-project.org")
install.packages('dplyr',repos = "http://cran.us.r-project.org")
install.packages("tidyr", repos = "http://cran.us.r-project.org")
install.packages('GGally',repos = "http://cran.us.r-project.org")
install.packages('memisc',repos = "http://cran.us.r-project.org")

library(ggplot2)
library(RColorBrewer)
library(readxl, quietly=TRUE)
library(gridExtra)
library(GGally)
library(readr)

```



```{r echo=FALSE, Load_the_Data}
# Load the Data
MyData <-read.csv("C:/eda-project/wineQualityWhites.csv")

```

## Univariate Plots Section -  and Preliminary exploration

```{r echo=FALSE}

# Running summaries to explore data 
# run summaries to understand the variables in the dataset
str(MyData)
# structure of dataset
print('')
summary(MyData)
#summary of data

```

MyData has 4898 observations and 13 variables.  11 of these varibles are numeric 
with 2 being integers.

```{r echo=FALSE}

# change the first column name from x to 'wines'
colnames(MyData)[1] <- "wines"
```


```{r echo=FALSE}

# to check for missing values
for (Var in names(MyData)) {
    missing <- sum(is.na(MyData[,Var]))
    if (missing > 0) {
        print(c(Var,missing))
    }
}
```

Confirms that MyData has no missing values.


```{r echo=FALSE}

#Distribution of fixed acidity 
library(ggplot2)
ggplot(aes(x = fixed.acidity), data = MyData) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = 3:16)

```

Fixed acidity is normally distributed at 6.8 g/dm^3 with some outliers.

```{r echo=FALSE}

# summary of fixed acidity
summary(MyData$fixed.acidity)

```

Fixed acidity has a normal distribution centred about 6.9.  The histogram shows 
that most white wines have an acidity ranging between 5.5 and 8.5.


```{r echo=FALSE}

#Distribution of volatile acidity 
ggplot(aes(x = volatile.acidity), data = MyData) +
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0.05, 1.12, 0.1))

```

Volatile acidity has a distribution slightly skewed to the right, with a centre 
of 0.26.  Most Wines have a VA between 0.15 and 0.35

```{r echo=FALSE}

# summary of fixed acidity
summary(MyData$volatile.acidity)

```


```{r echo=FALSE}

#Distribution of Citric acid 
ggplot(aes(x = citric.acid), data = MyData) +
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0, 1.8, 0.1))

```

```{r echo=FALSE}

summary(MyData$citric.acid)
```

Citric acid is normally distributed with most wines ranging between 0.2 to 0.5.  
Some very notable Outliers at 0.49 and 0.75. 


```{r echo=FALSE}

#Distribution of Residual sugar 
Plot1 <-  ggplot(aes(x = residual.sugar), data = MyData) +
  geom_histogram(binwidth = 0.1) 

Plot2 <-  Plot1 + scale_x_log10()
# transforming the data to increase visibility
 grid.arrange(Plot1, Plot2, ncol=2)
 ggsave('residual.sugar_hist.png')

  
```

```{r echo=FALSE}

summary(MyData$residual.sugar)

```

Residual sugar data is skewed to the right with the centre at 5.2g/dm^3 . To see
the distribution better we transform the data using Log 10 scale. This gives a 
bimodal distribution. The max sugar can be 65 but most wines have a value less 
than 15. Note the outlier near zero. 

```{r echo=FALSE}

#Distribution of chlorides 
ggplot(aes(x = chlorides), data = MyData) +
  geom_histogram(binwidth = 0.003) +
  scale_x_continuous(breaks = seq(0.01, 0.25, 0.02))

```

Chlorides have a normal distribution  with a slight skew to the right.  The 
centre is at 0.04300. Most wines fall between 0.025 - 0.065 mg/dm^3.

```{r echo=FALSE}

# Distribution of free sulphrdioxide
ggplot(aes(x = free.sulfur.dioxide), data = MyData) +
  geom_histogram(binwidth = 2) +
  scale_x_continuous(breaks = seq(0, 100, 10))

```

Free sulphur dioxide has a slight skew to the right with the centre being at 34.  
Most wines have a quantity less than 65 mg/dm^3.

```{r echo=FALSE}

# Distribution of total sulphurdioxide 
ggplot(aes(x = total.sulfur.dioxide), data = MyData) +
  geom_histogram(binwidth = 3) +
  scale_x_continuous(breaks = seq(5, 300, 20))

```

Total sulphur dioxide is normally distributed with a slight skew to the right. 
Most wines have a total between 65 and 215 mg/dm^3. The centre is at 134mg/dm^3.

```{r echo=FALSE}

# Distribution of density
ggplot(aes(x = density), data = MyData) +
  geom_histogram(binwidth = 0.0005) +
  scale_x_continuous(breaks = seq(0, 1.01, 0.005))

```

Density is normally distributed with a centre of 0.9940.  Most wines have a 
denstity between 0.9855 and 1.000. There some notable outliers.

```{r echo=FALSE}

# Distribution of pH
ggplot(aes(x = pH), data = MyData) +
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(2, 4.0, 0.1))
```

PH is symetric, With most wines ranging between 3.05 and 3.35.  The mean is at 
3.188.  This shows that most wines have a pH below 4 due to fermentation.

```{r echo=FALSE}

# Distribution of sulphates
ggplot(aes(x = sulphates), data = MyData) +
  geom_histogram(binwidth = 0.01) +
  scale_x_continuous(breaks = seq(0.2, 1.2, 0.1))

```

The distribution is slightly skewed to the right, the centre being at 0.47.  
Notable outliers at 0.36 and 0.5.  Most wines range between 0.35 - 0.6

```{r echo=FALSE}

# Distribution of alcohol
plot3 <- ggplot(aes(x = alcohol), data = MyData) +
  geom_histogram(binwidth = 0.1) +
  scale_x_continuous(breaks = seq(8.0, 16.0, 1))

plot4 <-  plot3 + scale_x_sqrt()
# transforming the data to increase visibility
 grid.arrange(plot3, plot4, ncol=2)
 ggsave('alcohol_hist.png')

```

There is a wide range in the alchol content of wines between 8.5 - 14 % by volume.
Alcohol data  has a bimodal distribution that seems to be skewed to the right 
with a major peak at 9.5.  The centre is at 10.40


```{r echo=FALSE}

#change quality from an int to a factor
MyData$quality <- factor(MyData$quality, ordered = TRUE)
is.factor(MyData$quality)

```

```{r echo=FALSE}
 
# Distribution of quality
ggplot(aes(x = quality), y = ..count../sum(..count..), data = MyData) +
  geom_bar()

```

```{r echo=FALSE}

# actual numbers in the bins
table(MyData$quality)
```
Majority of the wines are normal(5 - 7) than poor (3-4) or excellent one(8-9). 


```{r echo=FALSE}

# All the distributions of variables together - same scale
fixed.acidity <- qplot(data= MyData, x = fixed.acidity, binwidth = 0.1)
volatile.acidity <- qplot(data= MyData, x = volatile.acidity, binwidth = 0.01)
citric.acid <- qplot(data= MyData, x = citric.acid, binwidth = 0.01)
residual.sugar <- qplot(data= MyData, x = residual.sugar, binwidth = 0.5)
chlorides <- qplot(data= MyData, x = chlorides, binwidth = 0.003)
free.sulfur.dioxide <- qplot(data= MyData, x = free.sulfur.dioxide, binwidth = 2)
density <- qplot(data= MyData, x = density, binwidth = 0.0005)
pH <- qplot(data= MyData, x = pH, binwidth = 0.01)
sulphates <- qplot(data= MyData, x = sulphates, binwidth = 0.01)
alcohol <- qplot(data= MyData, x = alcohol, binwidth = 0.1)
quality <- qplot(data= MyData, x = quality)

grid.arrange(fixed.acidity,volatile.acidity,citric.acid,residual.sugar,chlorides,
     free.sulfur.dioxide, density, pH, sulphates,alcohol, quality, ncol=3)
ggsave('total_histograms.png')

```

```{r echo=FALSE}

# A correlation matrix plot 
library("GGally")
ggcorr(MyData, palette = "RdBu", hjust = 0.85, label = TRUE, label_size = 5, 
       label_color = "white")

```


### Observations:
The pairs showing a correlation of 0.3 and above(in red) are likely to have a 
significant effect on the quality of wine and shall 
be the subject of our investigation. Such are pairs are:
-fixed acidity & citric acid (0.3),
-residual sugar & free.sulphurdioxide(0.3),
-residual sugar & total.sulphurdioxide(0.4),
-residual sugar & density(0.8)
-chlorides & density(0.3)


### Univariate Analysis

The median quality is 6.  Majority of the wines are normal(5 - 7) than poor 
(3-4) or excellent one(8-9).
All the variables are normally distributed apart from residual sugar and alcohol
which are skewed to the right.  
Most wines have almost the same density and this varies between 0.9855 and 1.000.
Most wines have a pH of 3.4 or higher. About 75% of wines have quality that is 
lower than 6. The median percent alcohol content is 10.20 and the max percent 
alcohol content is 14.90.

#### What is the structure of your dataset?
The dataset - MyData has 4898 white wines(observations) with 13 features
(variables).  The column names of these variables are: fixed.acidity, 
volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide,
total.sulfur.dioxide, density, pH, sulphates, alcohol, and quality

#### What is/are the main feature(s) of interest in your dataset?
The main features of interest are quality, residual sugar, alcohol and acids - 
which I think play the biggest role in the quality of wine. The objective of the
analysis is to determine what features influence wine quality the most; then a 
predictive model of quality using these variables will be built.

#### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest? See the ggpairs correlation matrix.
All the pairs with correlation of 0.3 and above(in red) may have a significant
effect on the quality of wine and shall be investigated. such pairs are:
-fixed acidity & citric acid (0.3),
-residual sugar & free.sulphurdioxide(0.3),
-residual sugar & total.sulphurdioxide(0.4),
-residual sugar & density(0.8)

#### Did you create any new variables from existing variables in the dataset?
No, variables were created in this case. Quality which was a numeric variable 
was converted into a factor.

#### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this? 
Residual.sugar and alcohol are noted for having unusual distributions and 
outliers. Residual sugar has a long right tail meaning the data is skewed greatly 
by outliers - this  was then transformed in order to visualise the data better.
The alcohol distribution has a bimodal shape with a slight skew to the right. 



### Bivariate Plots Section
Features of interest are:
-fixed acidity & citric acid (0.3),
-residual sugar & free.sulphurdioxide(0.3),
-residual sugar & total.sulphurdioxide(0.4),
-residual sugar & density(0.8)
-chlorides & density(0.3)


```{r echo=FALSE, Bivariate_Plots} 

# exploring fixed acidity faceted with quality
# same bins and scale
qplot(x = fixed.acidity, data = MyData) + facet_wrap(~quality)

```

Fixed acidity has the same pattern like citric acid on quality.  9 and  3 have 
the least amount of citric acid while 6,5,7 respectively  have the highest. The 
wine documentation says, "found in small quantities, citric acid can add 
'freshness' and 'flavor'  and this varies between 0.3450 and 0.360 g / dm^3.


```{r echo=FALSE}
# citric acid vs quality

ggplot(aes(x = quality, y = citric.acid),
  data = MyData) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(0.2, 0.5)) + 
  stat_summary(geom = 'point', shape = 5, fun.y = mean)

```

 The y axis is limited to visualise the data better and eliminate outliers.  The
 mean citric acid for 3 and 9 are almost equal - aligns with the facet plot above.
 Based on the popular wines(5-7), 0.34 is an optimal value for citric acid.


```{r echo=FALSE}

# actual numbers - citric acid vs quality
by(MyData$citric.acid, MyData$quality, summary)

```
Increase in citric acid to a value of 0.36 favored the quality of wine

```{r echo=FALSE}

# residual sugar vs quality
# zoomed into the data by using a 95% quartile
# 95% also eliminates outliers 

ggplot(aes(x = quality, y = residual.sugar),
  data = MyData) + 
  geom_boxplot() + 
  ylim(0, quantile(MyData$residual.sugar, 0.95)) + 
  stat_summary(geom = 'point', shape = 5, fun.y = mean)
  
```
Apart from 5, the mean is pulled to the right by the outliers.  There appears no
clear trend of quality with residual sugar!

```{r echo=FALSE}
qplot(x = residual.sugar, data = MyData) + facet_wrap(~quality)
```

There is more sugar in 5, 6 and 7 and that could be partly the reason for their 
popularity but this same reason could account for why they're not the best quality.


```{r echo=FALSE}
# Actual numbers residual sugar vs quality
by(MyData$residual.sugar, MyData$quality, summary)
```

A decrease in sugar content improves quality


```{r echo=FALSE}

# Free sulphurdioxide vs quality
ggplot(aes(x = quality, y = free.sulfur.dioxide),
       data = MyData) + 
  geom_boxplot() + 
  coord_cartesian(ylim= c(0, 60)) + 
  stat_summary(geom = 'point', shape = 5, fun.y = mean)

```

The trend from 5-8 suggests that an increaste in free sulphurdioxide  favors 
quality.  Lets look at the values

```{r echo=FALSE}

#Get actual Numbers - free.sulfur.dioxide vs quality
by(MyData$free.sulfur.dioxide, MyData$quality, summary)

```

A decrease in free suphurdioxide from 35 to about 28 mg / dm^3 improves the quality

```{r echo=FALSE}

# total.sulphurdioxide vs quality
ggplot(aes(x = quality, y = total.sulfur.dioxide),
       data = MyData) + 
  geom_boxplot() + 
  coord_cartesian(ylim= c(70, 250)) + 
  stat_summary(geom = 'point', shape = 5, fun.y = mean)

```

The trend shows that about 125 mg / dm^3 of total sulphur dioxide are optimum for 
quality.  Let us get the exact values below.

```{r echo=FALSE}

# Actual Numbers - sulfurdioxide vs quality
by(MyData$total.sulfur.dioxide, MyData$quality, summary)

```

```{r echo=FALSE}

# Density vs quality
ggplot(aes(x = quality, y = density), data = MyData) + 
  geom_boxplot(color = 'black', fill = '#099DD9') + 
  coord_cartesian(ylim = c(0.9855, 1,1)) +
  stat_summary(geom = 'point', shape = 5, fun.y = mean) + 
  labs(x ="Quality score(0 -10)", y ='wine density g/cm^3') +
  ggtitle('Wine density by quality')

```

```{r echo=FALSE}

# Actual Numbers - density vs quality
by(MyData$density, MyData$quality, summary)

```

The trend  from 5 - 9 seems to show that lowering  density values(from 0.9953  
to 0.9903) seems to improve wine quality


```{r echo=FALSE}

# chlorides vs quality
ggplot(aes(x = quality, y = chlorides),
       data = MyData) + 
  geom_boxplot() + 
  coord_cartesian(ylim = c(0.01, 0.06)) +
  # limit the y axis to zoom into the data, eliminate outliers
  stat_summary(geom = 'point', shape = 5, fun.y = mean)

```

Lowering chloride levels from 0.046 to 0.031 improves the quality of wine


```{r echo=FALSE}

# actual Numbers - chlorides vs quality
by(MyData$chlorides, MyData$quality, summary)

```


```{r echo=FALSE}

# chlorides with a facet wrap
qplot(x = chlorides, data = MyData) + facet_wrap(~quality)

```

The trend  from a quality of 5 - 9 seems to show that lowering  chloride values 
from 0.04700  to 0.0310 g / dm^3 seems to improve wine quality


```{r echo=FALSE}

#Boxplot of  alcohol vs quality
ggplot(aes(x = quality, y = alcohol),
       data = MyData) + 
  geom_boxplot() +   
  coord_cartesian(ylim = c(9,13)) +
  # limit the y axis to zoom into the data, eliminate outliers
  stat_summary(geom = 'point', shape = 5, fun.y = mean)

```

Increase in alcohol content is in line with increased quality of wine


```{r echo=FALSE}

#Actual Numbers - alcohol vs quality
by(MyData$alcohol, MyData$quality, summary)

```
The values support the visualisation above


```{r echo=FALSE}

# Scatter plot of residual sugar vs density
ggplot(aes(x = residual.sugar, y = density),
  data = MyData) + 
  geom_point(alpha = 1/20) + 
  # add transparency to reduce overplotting
  scale_x_continuous(lim = c(0, 35), breaks=seq(0, 35, 4)) +
  scale_y_continuous(lim = c(0.985, 1.01), breaks=seq(0.985, 1.01, 0.01)) +
  # limit scale to zoom in
  geom_smooth(method="lm", color="red")

```

Density increases with increased sugar; the two variables are highly positively 
correlated at 0.84 

```{r echo=FALSE}

# Correlation of density and residual sugar
cor.test(MyData$density, MyData$residual.sugar, method = 'pearson')

```
```{r echo=FALSE}

# residual.sugar vs alcohol
ggplot(aes(x = residual.sugar, y = alcohol),
  data = MyData) +  
  geom_jitter(alpha = 1/10) +
  # add transparency to reduce overplotting
  # added some noise 
  scale_x_log10() + 
  geom_smooth(method="lm", color="red")

```

Increase in sugar does not lead to increase in alcohol. 


```{r echo=FALSE}

cor.test(MyData$residual.sugar, MyData$alcohol, method = 'pearson')

```
```{r echo=FALSE}

# Scatter plot of residual sugar vs total sulphates
ggplot(aes(x = total.sulfur.dioxide, y = residual.sugar),
  data = MyData) + 
  geom_point(alpha=1/10) +
  # add transparency to reduce overplotting
  scale_x_continuous(lim = c(0, 300), breaks=seq(0, 300, 20)) +
  scale_y_continuous(lim = c(0, 35), breaks=seq(0, 35, 4)) +
  # limit scale to zoom in
  geom_smooth(method="lm", color="red")

```
An increase in total sulphurdioxide shows an increase  in sugar with a fairly 
strong correlation of 0.4.  I wonder whether this a better determinant of wine 
quality!


```{r echo=FALSE}

cor.test(MyData$residual.sugar, MyData$total.sulfur.dioxide, method = 'pearson')

```

## Bivariate Analysis

#### Talk about some of the relationships you observed.
How did the feature(s) of interest vary with other features in \
the dataset?  
The question being addressed here is how quality varied in 
relation to fixed acidity, residual sugar, free.sulphurdioxide, 
total.sulphurdioxide, density and chloride.

Fixed acidity and citric acid show a 'U-like' pattern! starting from a high 
value of 0.345g/dm^3 with quality 3, the citric acid is lowered  to 0.32 and 
gradually increased to 0.36 at quality level 9.  The increase shows an in the 
quality.

Residual sugar -  is sugar left over after fermentation. There is no clear
trend here.  However Sugar content decreases from  7 at quality 5 to 2.2 g/dm^3
at 9. This is telling us that the best quality wine has less sugar.  
But again it depends on ones taste because quality 8 has a sugar level twice 
as much as that at nine. 

Free Sulphur dioxide: The wine quality from 5 - 9 shows a trend from 35 to 28 
mg/dm^3;showing that a decrease within this range improves quality. A similar 
trend is shown by the total sulphates since they both work as antioxidants and 
preservatives of wine.

Density:  a decrease in density from 0.9953 to 0.9903 corresponds to a quality 
increase of 5 to 9. It  makes sense since density results from a combination of 
alcohol and sugar.  And we see that density changes correspond to the changes 
both in alcohol and sugar. This is confirmed by the high positive correlation of 
density and sugar of 0.84.

An increase in alcohol content supports increased wine quality from 5 - 9.
While a decrease in the amount of chlorides which are essentially salts supports 
improved wine quality.


#### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)? Yee - its interesting to note that an 
increase in alcohol improves quality.  Note that the high correlation of residual 
sugar with density is due to the combination of both the residual sugar and alcohol.

#### What was the strongest relationship you found?
The strongest relationship is that of residual sugar and density which  gives a 
positive correlation of 0.84.

## Multivariate Plots Section

I would like to establish whether its density or total sulphurdioxide vs 
residual sugar that can better show the differences in wine quality.


```{r echo=FALSE}

# Scatter plot of residual sugar vs density vs quality
ggplot(aes(x = residual.sugar, y = density, color = quality), data = MyData) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter') + 
  # add transparency to reduce overplotting
  # add some noise with jitter
    scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_log10() +
  scale_y_log10(limits = c(0.985, 1.01), breaks=seq(0.985, 1.01, 0.01)) + 
  geom_smooth(method="lm", color="red") + 
  ggtitle("Density (log10) by Residual sugar (log10) and quality")

```
Interesting! - the red line clearly divides quality into 2 groups; we have wines
of quality 3-5 clearly above the line. 6 seems to lie in the boundary while 7,8 
and 9 are blow the line.  Density & residual sugar dynamics are a good choice
in determining wine quality.  I think this can make a good model.



```{r echo=FALSE}

# Scatter plot of residual sugar vs totl sulphates

ggplot(aes(x = residual.sugar, y = total.sulfur.dioxide, color = quality),
  data = MyData) + 
  geom_point(alpha = 0.5, size = 2, position = 'jitter') + 
  # add transparency to reduce overplotting
  # add some noise with jitter
   scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'quality', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_log10() +
  scale_y_log10() +
  ggtitle('Total sulphurdioxide (log10) by Residual sugar(log10) vs quality') +
    geom_smooth(method="lm", color="red")
```

The demarcation between the qualities is not clear at all, therefore, I would
not take this as a good model.


#### Building the linear model 
To create a linear model I will use the lm function. Density is the outcome and 
residual sugar is the predictor variable.  Note the use of the I Wrapper around 
each of the variables.

I can also update the previous model to add another factor  in the regression 
(using the syntax below). We can continue to make more complex models by adding 
more variables (chlorides, free.sulphurdioxide, fixed acidity, residual sugar 
etc).


```{r echo=FALSE}

library(memisc)

m1 <- lm(I(log(density)) ~ I(residual.sugar) , data = MyData)
m2 <- update(m1, ~ . + residual.sugar)
m3 <- update(m2, ~ . + total.sulfur.dioxide)
m4 <- update(m3, ~ . + fixed.acidity)
m5 <- update(m4, ~ . + chlorides)
m6 <- update(m5, ~ . + pH)
m7 <- update(m6, ~ . + alcohol)
m8 <- update(m7, ~ . + sulphates)
m9 <- update(m8, ~ . + citric.acid)

mod <- mtable(m1, m2, m3, m4, m5, m6, m7, m8, m9, sdigits = 8) 
# all are placed in a table with mtable()

mod
#print the mod
```
so much information, so many values here - better to get a summarised output!


```{r echo=FALSE}

# Examining a Fit
# print the m1 fit

summary(m1)
```
Based on the R2 score, residual sugar alone explains about 70.3% the variance in
quality

### Multivariate Analysis

#### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?
I investigated residual sugar vs density by quality and residual sugar vs 
total sulphurdioxide by  quality .  Of the two plots, residual sugar vs density
by quality provided the best demarcation within the groups of quality.

#### Were there any interesting or surprising interactions between features?
The plot of residual sugar vs total sulphurdioxide seemed interesting with a 
strong correlation of 0.4 however when combined with quality I was surprised 
that it did not show any clear trend.

#### OPTIONAL: Did you create any models with your dataset? Discuss the \
Yes - a Model was built for density vs residual sugar vs quality. Based on the 
R2 score, residual sugar alone explains about 70.3% of the variance in quality.
The high positive correlation between density and 
sugar in the matrix already hinted on this.

#### Strengths and limitations of your model.
Decision models are increasingly powerful for tasks requiring the impartial 
analysis of vast amounts of data. The beauty of this model is that it allows us 
to gather and process vast amounts of information/factors  in one place. The 
model also enables one to select the number of variables to use - in this case, 
I added 10 variables from the dataset. And such a model is not biases against 
anything like the recent wine it took or whether it got up on the wrong side of 
the bed.  This model weighs all data objectively and evenly.

The limitation is that it needs vast amounts of information and can not directly
influence the process to give the desired outcome. 

They can predict the quality of a wine vintage but won't make the wine any better,
reduce its acidity, improve the balance, or change the undertones. For these 
sorts of estimates, finding ways to avoid bias and maintain accuracy is essential.


### Final Plots and Summary

#### Plot One
```{r echo=FALSE}

#Distribution of Residual sugar 
Plot1 <-  ggplot(aes(x = residual.sugar), data = MyData) +
  geom_histogram(binwidth = 0.1) + 
  labs(x="Residual.sugar (g/dm^3)") + 
  ggtitle('Distribution of residual.sugar in g/dm^3')

Plot2 <-  Plot1 + scale_x_log10()
# transforming the data to increase visibility
 grid.arrange(Plot1, Plot2, ncol=2)
 ggsave('residual.sugar_hist.png')

  
```

#### Description One
Residual sugar data is skewed to the right with the centre at 5.2g/dm^3. Note the outliers especially about zero . To see the distribution better, the data is transformed using Log 10 scale. This gives a 
bimodal distribution as shown. The max sugar can be 65 but most wines have a value less 
than 15. Note the outlier near zero.


#### Plot Two
```{r echo=FALSE, Plot_Two}

# Density vs quality
ggplot(aes(x = quality, y = density), data = MyData) + 
  geom_boxplot(color = 'black', fill = '#099DD9') + 
  coord_cartesian(ylim = c(0.9855, 1,1)) +
  stat_summary(geom = 'point', shape = 5, fun.y = mean) + 
  labs(x ="Quality score(0 -10)", y ='wine density g/cm^3') +
  ggtitle('Wine density in g/cm^3 by quality')

```

#### Description Two

This is an interesting plot becuase  it shows  that the density of all wines is 
almost the same. Secondly it demonstrates how small minor changes  in density affect 
wine quality.

#### Plot Three
```{r echo=FALSE, Plot_Three}

# Scatter plot of residual sugar vs density vs quality
ggplot(aes(x = residual.sugar, y = density, color = quality), data = MyData) + 
  geom_point(alpha = 0.5, size = 3, position = 'jitter') + 
  # add transparency to reduce overplotting
  # add some noise with jitter
    scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Quality', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_log10() +
  scale_y_log10(limits = c(0.985, 1.01), breaks=seq(0.985, 1.01, 0.01)) + 
  geom_smooth(method="lm", color="red") + 
  labs(x="Residual.sugar (g/dm^3)", y ='wine density (g/cm^3)') +
  ggtitle("Density (log10) by Residual sugar (log10) and quality")

```


#### Description Three

This is an interesting plot! - the red line clearly divides quality into 2 groups;
The plot clearly shows the less density fines have better quality than the denser 
ones above the line.  Density & residual sugar dynamics are therefor a good choice
in determining wine quality.  I am confident that this makes a good model.

------

## Reflection

What a journey!  Its great to come to a close, I have certainly been challenged
and had fun too. I have had to reset my PC to factory standard twice to 
remove one drive  and delet it from the path of my downloads.  Debugging errors 
would sometimes take ages than was necessary. Other than that it was partly an
easy trip because the dataset is clean  and almost no data wrangling was  
necessary. Making the different plots, the adjustment, transformations and 
interpretion - has been a valuable experience.

The systematic exploration of the dataset from univariates to bivariates and 
finally multivariates helped me to progressively understand the dataset.
Uncovering insights especially when relating two variables and more was 
interesting. Now I have a good understanding of how the different variables 
in wine interact with each other to produce wine.  Its interesting to see how 
minor changes like with density or alchohol - can determine the quality of wine. 
Quality is subjective.  Striking a good balance is key.

What went well for me and what gave me the best feeling was when I built my model 
and understood that indeed the parameters I had zeroed in, one of them - residual 
sugar alone explains about 70.3% of the variance in quality. This was surprising!

I believe that there is always room for improvement.  This dataset can be made
richer by adding external variables like soil type, variety of the grape, weather
etc.  With that said, Wine quality is definately a very subjective measure.  
While beauty depends on the eyes of the beholder, wine depends on the palate of 
the individual!


